---
title: "Module5_Guided"
author: "Kip McCharen"
date: "7/22/2020"
output: pdf_document
---


***
  
## Question 1
  

```{r setup, include=FALSE}
library(faraway)
data(seatpos)
attach(seatpos)
MLR <- lm(hipcenter~., data=seatpos)
print(summary(MLR))
```


***
  
## Question 2
  
  
rejecting null over but not for t statistics, none of p values are significant
high R2 but not great for each one, probably multicollinearity
  
none of the predictors account for significant partial variance, multicollinearity
  
```{R}
preds <- cbind(Age, Weight, HtShoes, Ht, Seated, Arm, Thigh, Leg)
cor(preds)
```
  
A lot of high correlation, almost all values are above .5
  
```{R}
library(faraway)
vif(MLR)
```
  
Heights are of course correlated, other values all below 10, but still. 
  
```{r pressure, echo=FALSE}
round(cor(seatpos[,3:8]),3)
```

Correlation between Hts are almost 1. 

We know this group of predictors is similar, we can just remove one. 


***
  
## Question 3 
  


```{R}
pairs(preds)
```


***
  
## Question 4 
  

predictors good except age

```{r cars}
library(faraway)
vif(MLR)
```


***
  
## Question 5 
  

```{r pressure, echo=FALSE}
round(cor(seatpos[,3:8]),3)
```

***
  
## Question 6  
  
when peopl eare tall, limbs are longer 

Which value to keep? 
  height highly correlated with the rest
  probably an understanding piece of information for audiences

If we're automating seat position, we can easily use height. 

Height with shoes is more realistic. Just as high correlations as Height but you drive while wearing shoes. 

Thigh could resolve multicollinearity as much as possible. Least correlated with age and weight, while still highly correlated with height. 

***
  
## Question 7  
  
```{R}
MLR2 <- lm(formula = hipcenter ~Age+Weight+Ht)
vif(MLR2)
anova(MLR2,MLR)
```

High p value, should still reject the null. 

